Towards Practical First-Order Model Counting

### 2025-02-03 doc/paper/paper.tex
Section 4.2

The experiments were conducted using an Intel Skylake \SI{2.4}{\giga\hertz} CPU
with \SI{188}{\gibi\byte} of memory and CentOS~7. C++ programs were compiled
using the Intel C++ Compiler 2020u4. \textsc{FastWFOMC} ran on Julia~1.10.4,
while the other algorithms were executed on the Java Virtual Machine 1.8.0\_201.

We ran each algorithm on each benchmark using domains of sizes
$2^{1}, 2^{2}, 2^{3}$, and so on, until an algorithm failed to handle a domain
size due to timeout (of \SI{1}{\hour}), out of memory error, or out of precision
errors. While we separately measured compilation and inference time, we
primarily focus on total runtime, dominated by the latter.

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item Maybe add exact powers of two as vertical dotted lines (with both
          labels and lines in, say, red). Also try to expand this section a bit.
    \item We are not aware of any formulas on which \Cranetwo{} scales worse
          compared to either \textsc{ForcLift} or \textsc{FastWFOMC}. The one
          advantage that \textsc{FastWFOMC} has over \Cranetwo{} is its support
          for counting quantifiers.
    \item Regarding programming languages and accuracy, we verified that the
          answers match for smaller domain sizes. Also, although written in
          different languages, both \Cranetwo{} and \textsc{FastWFOMC} use the
          GNU Multiple Precision Arithmetic Library.
  \end{itemize}
}

\todo[inline]{Incorporate into the text below: It is worth remarking that
  \textsc{ForcLift} does not support arbitrary precision, and returns error for
  cases that requires arbitrary precision reasoning. }

@@

Another aspect of the experimental results that deserves separate discussion is
compilation. Both Julia and Scala use just-in-time (JIT) compilation, which
means that \textsc{FastWFOMC} and \textsc{ForcLift} take longer to run on the
smallest domain size, where most JIT compilation occurs. In the case of
\Cranetwo{}, it is only run once per benchmark, so the JIT compilation time is
included in its overall runtime across all domain sizes. Additionally, while
\textsc{ForcLift}'s compilation is generally faster than that of \Cranetwo{},
neither significantly affects overall runtime. Specifically, \textsc{ForcLift}
compilation typically takes around \SI{0.5}{\second}, while \Cranetwo{}
compilation takes around \SI{2.3}{\second}.

@@

>>>>

The experiments were conducted using an Intel Skylake \SI{2.4}{\giga\hertz} CPU
with \SI{188}{\gibi\byte} of memory and CentOS~7. C++ programs were compiled
using the Intel C++ Compiler 2020u4. \textsc{FastWFOMC} ran on Julia~1.10.4,
while the other algorithms were executed on the Java Virtual Machine 1.8.0\_201.
Note that, although implemented in different languages, both \Cranetwo{} and
\textsc{FastWFOMC} use the GNU Multiple Precision Arithmetic Library for
arbitrary-precision arithmetic.

We ran each algorithm on each benchmark using domains of sizes
$2^{1}, 2^{2}, 2^{3}$, and so on, until an algorithm failed to handle a domain
size due to timeout (of \SI{1}{\hour}), out of memory error, or out of precision
errors. While we separately measured compilation and inference time, we
primarily focus on total runtime, dominated by the latter. We verified the
accuracy of the numerical answers using the corresponding integer sequences on
the On-Line Encyclopedia of Integer Sequences~\cite{oeis}.

@@

@@

One might notice that the runtime of \textsc{FastWFOMC} and \textsc{ForcLift} is
slightly higher on the smallest domain size. This peculiarity is the consequence
of \emph{just-in-time} (JIT) compilation. As \Cranetwo{} is only run once per
benchmark, the JIT compilation time is included in its overall runtime across
all domain sizes. Additionally, while \textsc{ForcLift}'s compilation is
generally faster than that of \Cranetwo{}, neither significantly affects overall
runtime. Specifically, \textsc{ForcLift} compilation typically takes around
\SI{0.5}{\second}, while \Cranetwo{} compilation takes around \SI{2.3}{\second}.

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item Somewhere: reference the appendix, saying that `we provide a more
          detailed comparison of the capabilities of the algorithms'
  \end{itemize}
}
